---
title: "Time varying price examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time varying price examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618
)
```

```{r packages, message=FALSE, warning=TRUE}
library(ggplot2)
library(dplyr)
library(pinkescape)
```

## Implementing time-varying changes in prices

To implement changes in prices over time, we can modify the three following arguments. If either `price_linear_change` or `price_cv` is set to non-zero values, then dynamic prices are turned on in the simulation framework.

```{r eval=FALSE, echo=TRUE}
sim(...,
    price_linear_change = 0,
    price_acf = 0.7, 
    price_cv = 0)
```

We can demonstrate this by first running a scenario as a baseline,

```{r eval=TRUE, echo=TRUE}
df_baseline = sim(rec_std = 0.05)
```

And then we can add 3 scenarios for comparison. First, we'll simulate the effects of a deterministic price decline of -0.01. This is the drift in log space, so that

$$
log(p_{t}) = log(p_{t-1}) + u
$$

such that $u = -0.01$.

```{r}
df_det = sim(rec_std = 0.05, price_linear_change = -0.01)
```

Second, we'll fit a stochastic model without drift ($u = 0$). For this model, 

$$
log(p_{t}) = log(p_{t-1}) + \delta_t,  \\ \delta_t \sim N(0,\sigma)
$$

```{r}
df_stoch = sim(rec_std = 0.05, price_cv = 0.1)
```

Third, we'll fit a model that combines the stochastic variation with drift. 

```{r}
df_comb = sim(rec_std = 0.05, price_cv = 0.1, price_linear_change = -0.01)
```


### Summarizing results

We want to summarize the total net benefits, and estimates of recruitment across these scenarios. 

First, we can combine all of the individual results in a dataframe.

```{r}
df_baseline$scenario = "Baseline"
df_det$scenario = "Drift"
df_stoch$scenario = "Stochastic"
df_comb$scenario = "Drift + Stochastic"

df = rbind(df_baseline, df_det, df_stoch, df_comb)
```

We can look at total net benefits across scenario. This is shown in log-space because the stochastic models have a lot of uncertainty,

```{r}
df_summary = mutate(df,
                              net_ben = discount*net_benefits) |>
  group_by(scenario, sim) |>
  summarise(tot_ben = sum(net_ben))

ggplot(df_summary, aes(scenario, log(tot_ben),group=scenario)) +
  geom_boxplot()
```


We could also look at making time series of benefits over time for each scenario. To do this, we can summarize across time slices,

```{r}
df_summary = mutate(df,
                              net_ben = discount*net_benefits) |>
  group_by(scenario, t) |>
  summarise(mean_ben = median(net_ben),
            lo = quantile(net_ben,0.025),
            hi = quantile(net_ben,0.975))

ggplot(df_summary, aes(t, mean_ben)) +
  geom_ribbon(aes(ymin=lo, ymax = hi)) + 
  geom_line() + 
  facet_wrap(~scenario)
```

