---
title: "Simulation examples"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618
)
```

```{r packages, message=FALSE, warning=TRUE}
library(ggplot2)
library(dplyr)
library(pinkescape)
```

## Comparing variability of recruitment on net benefits

We will simulate a low and high variance scenario, calculate net benefits for each simulation and plot results. 
```{r}
df_lo = sim(rec_std = 0.05)
df_hi = sim(rec_std = 0.2)

# summarize total benefits
df_lo_summary = mutate(df_lo,
                              net_ben = discount*net_benefits) %>%
  group_by(sim) %>%
  summarise(tot_ben = sum(net_ben))

df_hi_summary = mutate(df_hi,
                              net_ben = discount*net_benefits) %>%
  group_by(sim) %>%
  summarise(tot_ben = sum(net_ben))

df_lo_summary$sd = 0.05
df_hi_summary$sd = 0.2
ggplot(rbind(df_lo_summary,df_hi_summary), aes(sd, tot_ben,group=sd)) +
  geom_boxplot()

```

## Comparing effect of recruitment autocorrelation on net benefits

Similar to above, we can compare a white noise to autocorrelated scenario,  

```{r}
df_lo = sim(rec_acf = 0)
df_hi = sim(rec_acf = 0.7)
# summarize
df_lo_summary = mutate(df_lo,
                              net_ben = discount*net_benefits) %>%
  group_by(sim) %>%
  summarise(tot_ben = sum(net_ben))
df_hi_summary = mutate(df_hi,
                              net_ben = discount*net_benefits) %>%
  group_by(sim) %>%
  summarise(tot_ben = sum(net_ben))
df_lo_summary$acf = 0
df_hi_summary$acf = 0.7
ggplot(rbind(df_lo_summary,df_hi_summary), aes(acf, tot_ben,group=acf)) +
  geom_boxplot()

```

## Adding harvest variability 

We can introduce variability in realized harvest by specifying the `harvest_CV` parameter, which is a lognormal variable multiplied by expected harvest. Realistic values are probably in the 0.5 - 1.0 range.

```{r eval=FALSE}
sim(harvest_CV = 0.5)
```


## Using custom Ricker parameters

Instead of the defaults, we can create a dataframe with custom parameters and pass these in. The default parameters can be found using the `get_ricker()` function,  

```{r}
pars <- get_ricker()
```

But let's change the Ricker $a$ and $b$ parameters. We have a custom 

```{r}
pars$a = c(1,1,2,2)
pars$b = runif(4)
```

These can be passed into the simulation now

```{r}
custom_sims <- sim(ricker_pars = pars)
```


