---
title: "Simulation examples"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette: default
vignette: |
  %\VignetteIndexEntry{Simulation examples} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.asp = 0.618
)
```

```{r packages, message=FALSE, warning=TRUE}
library(ggplot2)
library(dplyr)
library(pinkescape)
library(kableExtra)
```

## Using custom Ricker parameters

Instead of the defaults, we can create a dataframe with custom parameters and pass these in. The default parameters can be found using the `get_ricker()` function,  

```{r}
pars <- get_ricker()
```

But let's change the Ricker $a$ and $b$ parameters. We have a custom 

```{r}
pars$a = c(2.14622314, 1.09373485, 1.20498185, 2.64594179)
pars$b = c(-0.16520772,-0.04008856,-0.09372470,-0.47663675)
pars$S_star=c(6.12339,13.4478,6.88375,2.74133)
pars$real_price = c(2099586.443,2099586.443,2132653.372,2132653.372)
pars$cst_param_calib=c(13140389.91,13140389.91,9435607.205,9435607.205)
pars$S_star_deterministic = c(6.38142, 10.9406, 6.78888, 3.02635)
```

Create a data frame for passing arguments to simulations.  

```{r}
runs_det = expand.grid(run = c("odd","even"),
                   escapement_rule = c("both","pre","post"),
                   rec_std = c(0.02, 0.2),
                   time_lag = c(0,10),
                   deterministic_model = TRUE)
runs_det$name = "SW"
runs_det$name[which(runs_det$escapement_rule=="pre")] = "PR"
runs_det$name[which(runs_det$escapement_rule=="post")] = "PS"

runs_stoch = expand.grid(run = c("odd","even"),
                   escapement_rule = c("both","pre","post"),
                   rec_std = c(0.02, 0.2),
                   time_lag = c(0,10),
                   deterministic_model = FALSE)
runs_stoch$name = "ST-SW"
runs_stoch$name[which(runs_stoch$escapement_rule=="pre")] = "ST-PR"
runs_stoch$name[which(runs_stoch$escapement_rule=="post")] = "ST-PS"

runs = rbind(runs_det, runs_stoch)
```

Loop through and bind the simulations together

```{r}
for(i in 1:nrow(runs)) {
  s = sim(rec_std = runs$rec_std[i],
          ricker_pars = pars,
          pr_12 = 0.118,
          pr_21 = 0.202,
          run=runs$run[i], 
          time_lag = runs$time_lag[i], 
          escapement_rule = runs$escapement_rule[i])
  # add extra info
  s$id = i # just the row in the runs dataframe
  s$rec_std = runs$rec_std[i]
  s$run = runs$run[i]
  s$escapement_rule = runs$escapement_rule[i]
  s$name = runs$name[i]
  s$time_lag = runs$time_lag[i]
  # save output to single dataframe
  if(i == 1) {
    output = s
  } else {
    output = rbind(output, s)
  }
}
```

For MSY simulations, we can do something similar and just change S_star 
```{r}
pars$S_star=c(6.05299, 24.9448,10.6695, 2.09803)
```

```{r}
runs_msy = runs_det
runs_msy$name = paste0("MSY-",runs_msy$name)
  
for(i in 1:nrow(runs_msy)) {
  s = sim(rec_std = runs_msy$rec_std[i],
          ricker_pars = pars,
          pr_12 = 0.118,
          pr_21 = 0.202,
          run=runs_msy$run[i], 
          time_lag = 10, 
          escapement_rule = runs_msy$escapement_rule[i])
  # add extra info
  s$id = i + nrow(runs)# just the row in the runs dataframe
  s$rec_std = runs_msy$rec_std[i]
  s$run = runs_msy$run[i]
  s$escapement_rule = runs_msy$escapement_rule[i]
  s$name = runs_msy$name[i]
  s$time_lag = runs_msy$time_lag[i]
  # save output to single dataframe
  if(i == 1) {
    output_msy = s
  } else {
    output_msy = rbind(output_msy, s)
  }
}
```

Bind output together

```{r}
output = rbind(output, output_msy)
output$rec_std = as.factor(output$rec_std)
output$name = as.factor(output$name)
```

Summarize 

```{r}
out = dplyr::group_by(output, id) %>%
  dplyr::summarise(total_ben = sum(net_benefits/1.0e6),
                   ben_mean = mean(net_benefits/1.0e6,0.975),
                   ben_hi = quantile(net_benefits/1.0e6,0.975),
                   ben_lo = quantile(net_benefits/1.0e6,0.025),
                   spawn_mean=mean(spawners),
                   spawn_hi = quantile(spawners,0.975),
                   spawn_lo = quantile(spawners,0.025),
                   rec_mean=mean(rec),
                   rec_hi = quantile(rec,0.975),
                   rec_lo = quantile(rec,0.025),
                   harv_mean=mean(harvest),
                   harv_hi = quantile(harvest,0.975),
                   harv_lo = quantile(harvest,0.025),
                   spawn_sd = sd(spawners),
                   harv_sd = sd(harvest),
                   rec_std = rec_std[1],
                   run = run[1],
                   escapement_rule = escapement_rule[1],
                   name = name[1],
                   time_lag = time_lag[1])
out$Level = paste0("Rec_sigma: ", out$rec_std,", Lag: ", out$time_lag)
```

### Plots

```{r}
p1 = ggplot(out, aes(name,ben_mean,fill=Level,col=Level)) + 
  geom_point(position = position_dodge2(0.4)) + 
  geom_linerange(aes(ymin = ben_lo, ymax = ben_hi),
                 position = position_dodge2(0.4)) + 
  xlab("") + ylab("Net Benefits (Million $)") + 
  facet_wrap(~run, scale="free",ncol = 1) +
  theme_bw()
p1
ggsave(p1, file="net_benefits.jpeg", height=5, width=7)
```

```{r}
p2 = ggplot(out, aes(name,rec_mean,fill=Level,col=Level)) + 
  geom_point(position = position_dodge2(0.4)) + 
  geom_linerange(aes(ymin = rec_lo, ymax = rec_hi),
                 position = position_dodge2(0.4)) + 
  xlab("") + ylab("Recruitment") + 
  facet_wrap(~run, scale="free",ncol = 1) +
  theme_bw()
p2
ggsave(p2, file="recruitment.jpeg", height=5, width=7)
```

```{r}
p3 = ggplot(out, aes(name,harv_mean,fill=Level,col=Level)) + 
  geom_point(position = position_dodge2(0.4)) + 
  geom_linerange(aes(ymin = harv_lo, ymax = harv_hi),
                 position = position_dodge2(0.4)) + 
  xlab("") + ylab("Harvest") + 
  facet_wrap(~run, scale="free",ncol = 1) +
  theme_bw()
p3
ggsave(p3, file="harvest.jpeg", height=5, width=7)
```
