---
title: "Hidden Markov Models for environmental covariates (Ohlberger et al.)"
output:
  word_document: default
  pdf_document: default
---

## Approach

Using the covariates in Ohlberger et al., we can fit some simple HMMs to the environmental data to identify regimes and estimate transition probabilities. 

## SST - only

Jan's model included spring SST anomalies in the easter Gulf of Alaska, and those deviations can be included as an initial response to be modeled.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
max_iter = 10
```

```{r echo=FALSE}
library(ggplot2)
library(depmixS4)
d = read.csv("data.csv")

# fit HMM to SST data alone
df = data.frame(year = d$Year, y = d$egoa.spr.sst)

  # options for family are binomial, Gamma, gaussian, poisson, multinomial
mod <- depmix(list(y ~ 1), 
                family = list(gaussian()),
                data = df, nstates = 2, 
                em.control = list(maxit=2000, tol=1e-12))
  
fm <- fit(mod, verbose = FALSE) 
#get the estimated state for each timestep 
post_states <- depmixS4::posterior(fm)

df$states = as.character(post_states[,1])
ggplot(df, aes(year, y)) + 
  geom_rect(aes(xmin = year - 0.5, xmax = year + 0.5, 
                ymin = -Inf, ymax = Inf, fill = states), 
            alpha = 0.3) +
  geom_line(color = "grey20", size = 0.8) + 
  #geom_point(color = "grey20", size = 2) + 
  theme_bw() + 
  scale_fill_viridis_d(option = "magma", begin = 0.3, end = 0.8) +
  xlab("") + 
  ylab("Spring SST anomaly (GOA)") +
  labs(fill = "State")
ggsave("HMM_figure.png", width=7, height=6)
```


```{r}
# df$state = depmixS4::posterior(fm)[,1]
# 
# ggplot(df, aes(year, state)) + geom_line() + geom_point() + 
#   theme_bw() + xlab("") + ylab("Estimated state")

```

```{r}
table(post_states$state)
```

We can also extract the transition matrix, which is 

```{r}
m = matrix(0,2,2)
rownames(m) = c("From 1", "From 2")
colnames(m) = c("To 1", "To 2")
m[1,] = fm@transition[[1]]@parameters$coefficients
m[2,] = fm@transition[[2]]@parameters$coefficients
knitr::kable(m,digits = 3)
```



We can also extract the transition matrix, which is 

```{r}
m = matrix(0,2,2)
rownames(m) = c("From 1", "From 2")
colnames(m) = c("To 1", "To 2")
m[1,] = fm@transition[[1]]@parameters$coefficients
m[2,] = fm@transition[[2]]@parameters$coefficients
knitr::kable(m,digits = 3)
```

\break

### Notes:

During our call on 12/10, we discussed a couple options for extending these models to make them more useful. Specifically, we discussed modeling the following responses: 

* Harvests
* Run Size
* Productivity, ln(R/S)

For each of the above, we also could include hatchery releases as a potential driver (lags discussed below).

### Harvest models

```{r}
ggplot(d, aes(Year, log(PWS_pink_harvests))) + geom_line() + geom_point() +
  theme_bw() + xlab("") + ylab("PWS harvest, ln(millions)")
```

Because this increases so fast early in the time series, we might want to restrict the model to years 1985 - present. The challenge here is that the model wants to assign all recent years to a single state, which isn't very interesting. The choice of year doesn't really affect this (though when only data from the mid-1990s onward is used, the HMM just filters out high vs low pink years, so is oscillating back and forth nearly every year). 

```{r results="hide"}

sub = dplyr::filter(d,Year>=1985, Year <= 2018)

sub$covar = log(sub$PWS_released_prev_Total)

models = expand.grid(mean_cov = c("Year","1"),
                     trans_cov = c("Year","1"))
models$aic = 0
best_aic = 1.0e10

for(i in 1:nrow(models)) {
  # specifications of how to model,
# coming from somewhere else
outcome <- "log(PWS_pink_harvests)"

f1 <- as.formula(
  paste(outcome, 
        paste(models$mean_cov[i], collapse = " + "), 
        sep = " ~ "))
f2 <- as.formula(
  paste("",
        paste(models$trans_cov[i], collapse = " + "), 
        sep = " ~ "))

  mod <- depmix(list(f1), 
              transition = f2,
                family = list(gaussian()),
                data = sub, nstates = 2, 
                em.control = list(maxit=2000, tol=1e-12))

    best_aic = 1.0e10
  for(iter in 1:max_iter) {
    set.seed(iter ^ 2)
    fm <- fit(mod, verbose = FALSE)
    if(AIC(fm) < best_aic) {
      best_aic = AIC(fm)
      best_model = fm
    }
  } 
  models$aic[i] = AIC(fm)
  if(models$aic[i] < best_aic) {
    best_model = fm
    best_aic = models$aic[i]
  }
}


#get the estimated state for each timestep 
post_states <- depmixS4::posterior(best_model)

sub$state = depmixS4::posterior(best_model)[,1]

```
```{r}

sub$state = as.factor(sub$state)
ggplot(sub, aes(Year, log(PWS_pink_harvests), col = state, group=NA)) + geom_line() + geom_point() + 
  theme_bw() + xlab("") + ylab("log(PWS_pink_harvests)")

```


Second, we can try to model this as driven by pink salmon releases. That time series looks like this. The releases will be lagged one year here. 

```{r}
ggplot(dplyr::filter(d,Year<2020), aes(Year, PWS_released_Total)) + geom_line() + geom_point() +
  theme_bw() + xlab("") + ylab("PWS releases, (millions)")
```

Including the releases as a driver doesn't seem to do much, but letting the mean of each state be linearly changing over time lets the model separate out high/low pink salmon years.

```{r results="hide"}
sub = dplyr::filter(d,Year>=1985, Year <= 2018)

sub$covar = log(sub$PWS_released_prev_Total)

models = expand.grid(mean_cov = c("Year","covar","1","Brood.Line"),
                     trans_cov = c("Year","covar","1","Brood.Line"))
models$aic = 0
best_aic = 1.0e10
for(i in 1:nrow(models)) {
  # specifications of how to model,
# coming from somewhere else
outcome <- "log(PWS_pink_harvests)"

f1 <- as.formula(
  paste(outcome, 
        paste(models$mean_cov[i], collapse = " + "), 
        sep = " ~ "))
f2 <- as.formula(
  paste("",
        paste(models$trans_cov[i], collapse = " + "), 
        sep = " ~ "))

  mod <- depmix(list(f1), 
              transition = f2,
                family = list(gaussian()),
                data = sub, nstates = 2, 
                em.control = list(maxit=2000, tol=1e-12))
  this_aic = 1.0e10
  for(iter in 1:max_iter) {
    set.seed(iter ^ 2)
    fm <- try(fit(mod, verbose = FALSE), silent = TRUE)
    if(class(fm)!="try-error") {
      if(AIC(fm) < this_aic) {
        this_aic = AIC(fm)
        fm_model = fm
      }
    }
  }
  
  models$aic[i] = AIC(fm_model)
  if(models$aic[i] < best_aic) {
    best_model = fm
    best_aic = models$aic[i]
  }
}

#get the estimated state for each timestep 
post_states <- depmixS4::posterior(best_model)

sub$state = as.factor(depmixS4::posterior(best_model)[,1])

```

```{r}
ggplot(sub, aes(Year, log(PWS_pink_harvests), col=state, group=NA)) + geom_line() + geom_point() + 
  theme_bw() + xlab("") + ylab("") + ggtitle("log(PWS_pink_harvests)")
```

```{r}
knitr::kable(models, digits=3)
```

\break  

### Run size models

```{r results = "hide"}
sub = dplyr::filter(d,Year>=1985, Year <= 2018)

sub$covar = log(sub$PWS_released_prev_Total)

models = expand.grid(mean_cov = c("Year","covar","1","Brood.Line"),
                     trans_cov = c("Year","covar","1","Brood.Line"))
models$aic = 0
best_aic = 1.0e10
for(i in 1:nrow(models)) {
  # specifications of how to model,
# coming from somewhere else
outcome <- "log(Run.Size)"

f1 <- as.formula(
  paste(outcome, 
        paste(models$mean_cov[i], collapse = " + "), 
        sep = " ~ "))
f2 <- as.formula(
  paste("",
        paste(models$trans_cov[i], collapse = " + "), 
        sep = " ~ "))

  mod <- depmix(list(f1), 
              transition = f2,
                family = list(gaussian()),
                data = sub, nstates = 2, 
                em.control = list(maxit=2000, tol=1e-12))

  this_aic = 1.0e10
  for(iter in 1:max_iter) {
    set.seed(iter ^ 2)
    fm <- try(fit(mod, verbose = FALSE), silent = TRUE)
    if(class(fm)!="try-error") {
      if(AIC(fm) < this_aic) {
        this_aic = AIC(fm)
        fm_model = fm
      }
    }
  }
  
  models$aic[i] = AIC(fm_model)
  if(models$aic[i] < best_aic) {
    best_model = fm
    best_aic = models$aic[i]
  }
}


#get the estimated state for each timestep 
post_states <- depmixS4::posterior(best_model)

sub$state = as.factor(depmixS4::posterior(best_model)[,1])
```

Here, the best model is a model that lets the means (state 1/2) vary by brood line. And hatchery releases are included as a covariate on the transition matrix so that the probability of switching is also increasing over time (as a function of releases).  

```{r}
ggplot(sub, aes(Year, log(Run.Size), col = state, group=NA)) + geom_line() + geom_point() + 
  theme_bw() + xlab("") + ylab("log(Run.Size)") + ggtitle ("Run size models")
```

```{r}
knitr::kable(models, digits=3)
```

\break  

### Productivity models

```{r results = "hide"}
sub = dplyr::filter(d,Year>=1985, Year <= 2018)

sub$covar = log(sub$PWS_released_prev_Total)

models = expand.grid(mean_cov = c("Year","covar","1", "Brood.Line"),
                     trans_cov = c("Year","covar","1", "Brood.Line"))
models$aic = 0
best_aic = 1.0e10
for(i in 1:nrow(models)) {
  # specifications of how to model,
# coming from somewhere else
outcome <- "lnRS"

f1 <- as.formula(
  paste(outcome, 
        paste(models$mean_cov[i], collapse = " + "), 
        sep = " ~ "))
f2 <- as.formula(
  paste("",
        paste(models$trans_cov[i], collapse = " + "), 
        sep = " ~ "))

  mod <- depmix(list(f1), 
              transition = f2,
                family = list(gaussian()),
                data = sub, nstates = 2, 
                em.control = list(maxit=2000, tol=1e-12))

  this_aic = 1.0e10
  for(iter in 1:max_iter) {
    set.seed(iter ^ 2)
    fm <- try(fit(mod, verbose = FALSE), silent = TRUE)
    if(class(fm)!="try-error") {
      if(AIC(fm) < this_aic) {
        this_aic = AIC(fm)
        fm_model = fm
      }
    }
  }
  
  models$aic[i] = AIC(fm_model)
  if(models$aic[i] < best_aic) {
    best_model = fm
    best_aic = models$aic[i]
  }
}


#get the estimated state for each timestep 
post_states <- depmixS4::posterior(best_model)

sub$state = as.factor(depmixS4::posterior(best_model)[,1])

ggplot(sub, aes(Year, lnRS, col = state, group=NA)) + geom_line() + geom_point() + 
  theme_bw() + xlab("") + ylab("lnRS") + ggtitle ("Productivity models")

```

The best model of productivity includes hatchery releases as a predictor of the state means -- but not transition probabilities.  

```{r}
knitr::kable(models, digits=3)
```

